### 2017-05-02 R Code/Simulation Study/writeUp.Rmd

---
title: "Simulation Study for Statistical Data Science Conference"
author: "J.P. Meagher"
date: "1 May 2017"
output: html_document
---

The object of this simulation study is to demonstrate the conditions required for type II maximum likelihood estimation to successfully estimate hyperparameters for a Phylogenetic Ornstein-Uhlenbeck Process.

The Phylogenetic Ornsetin-Uhlenbeck Process kernel is
\[
k(x,x') = \sigma_p \exp \left( \frac{|x - x'|}{\ell} \right)  + \sigma_n
\]
where \(\sigma_p\) is the phylogenetic noise, \(\ell\) the phylogenetic length-scale, and \(sigma_n\) the non-phylogenetic noise.

The key parameter here is \(\ell\). It can be easily shown that when \(\ell >> |x - x'|\) there is a high degree of correlation between process observations at \(x\) and \(x'\). When \(\ell << |x - x'|\) there is little correlation and the process becomes indistinguishable from white noise. With this in mind fix \(\sigma_p = 1\) and \(\sigma_n = 1\).

Investigate the accuracy for the process over the tree for \(\ell = \{10, 25, 50, 100, 200\}\) considering \(n = \{1, 2, 4, 8\}\) observations at each tip of the tree

```{r}
library(sdsBAT)

n <- 1
l <- 10

tree <- expand_tree(phylogeny, n)
sim_10_1 <- pou_simulation_study(log(c(1,l,1)), tree, 100, logl_function = pou_logl_fast, optim_function = 'uobyqa', upper_initialisation = c(2, 150, 2), n_restarts = 5)

pou_simulation_study

pou_simulation_study(log(c(1,10,1)), phylogeny, 10, n_restarts = 5)

```