### 2017-05-05 R Code/Simulation Study/writeUp.Rmd
Simulated Hyperparameters and Component code

---
title: "Simulation Study for Statistical Data Science Conference"
author: "J.P. Meagher"
date: "1 May 2017"
output: html_document
---

The object of this simulation study is to demonstrate the conditions required for type II maximum likelihood estimation to successfully estimate hyperparameters for a Phylogenetic Ornstein-Uhlenbeck Process.

The Phylogenetic Ornsetin-Uhlenbeck Process kernel is
\[
k(x,x') = \sigma_p \exp \left( \frac{|x - x'|}{\ell} \right)  + \sigma_n
\]
where \(\sigma_p\) is the phylogenetic noise, \(\ell\) the phylogenetic length-scale, and \(sigma_n\) the non-phylogenetic noise.

The key parameter here is \(\ell\). It can be easily shown that when \(\ell >> |x - x'|\) there is a high degree of correlation between process observations at \(x\) and \(x'\). When \(\ell << |x - x'|\) there is little correlation and the process becomes indistinguishable from white noise. With this in mind fix \(\sigma_p = 1\) and \(\sigma_n = 1\).

Investigate the accuracy for the process over the tree for \(\ell = \{10, 25, 50, 100, 200\}\) considering \(n = \{1, 2, 4, 8\}\) observations at each tip of the tree.

Create a dataframe storing the results of 1000 simulations for the vlaues of \(\ell, n \) of interest.

Note that `uobyqa()` offers the fastest optimisation method, `'Nelder-Mead'` and `'BFGS'` are roughly comparable to each other and only slightly slower, while `'CG'` is far slower. I will use `uobyqa` with 5 restarts.

5 samples at each valeue take about 23 seconds, should take just over an hour to get the 1000 samples I want.

```{r, cache=TRUE}
library(sdsBAT)
set.seed(7)

n <- c(1, 2, 4, 8)
l <- c(10, 25, 50, 100, 200)
N <- 1000L

dummy <- matrix(1:20, nrow = length(n), ncol = length(l))
rownames(dummy) <- c('n.1', 'n.2', 'n.4', 'n.8'); colnames(dummy) <- c('l.10', 'l.25', 'l.50', 'l.100', 'l.200')

simulations <- data.frame(dummy)
for(i in 1:length(n)){
  for(j in 1:length(l)){
    tree <- expand_tree(phylogeny, n[i])
    experiment <- suppressMessages(pou_simulation_study(log(c(1,l[j],1)), tree, N, logl_function = pou_logl_fast, optim_function = 'uobyqa', upper_initialisation = c(2, 150, 2), n_restarts = 5))
    simulations[[j]][[i]] <- list(experiment)
  }
}

panel.hist <- function(x, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(usr[1:2], 0, 1.5) )
    h <- hist(x, plot = FALSE)
    breaks <- h$breaks; nB <- length(breaks)
    y <- h$counts; y <- y/max(y)
    rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}
```